Cover letter for the original IJRR submission

### 2017-12-13 review_tro_2/reviewer_response.tex

\section{Answers to the associate editor}

\quot {The C++/Java-style code listings are unnecessary and do not conform
to journal standards.  The authors rewrite their pseudocode to use
rigorous mathematical notation.  Suggest replacing "object.attribute"
with "Attribute(object)", and describing the attributes of the objects
in the text.
}

\as{We rewrote Algorithm 1, 2 and 3 accordingly to the recommendations of the Associate Editor, which we thank for the comment that allowed us to spot
a typo in pseudocode. Furthermore, it also allowed us to reduce the length of the pseudocode.}
\leavevmode 
\quot {The URLs in the paper seem rather brittle and are not suitable for
archival purposes.  Two URLs point to specific branches of various
Github repositories, and another refers to the documentation of a
Blender tool.  The authors should create a static web page for their
software which is guaranteed to persist for years to come... }

\as{Thanks for the comment. In fact, the github urls are not pointing to branches, but to release versions of the code, presented as zip files, which make them permanent
links. However, we agree with the Associate Editor that regrouping all the data and information on a single web page is relevant. So, we uploaded the source code to the first author
website (\url{http://stevetonneau.fr/files/publications/isrr15/tro_install.html}), and provided all the required information to install the material in it. The user has the choice to download the code either from the github permanent link or directly from the author's website.}

### 2017-12-13 review_tro_2/reviewer_response.tex

\as{Thanks for the comment. In fact, the github urls are not pointing to branches, but to release versions of the code, presented as zip files, which make them permanent
links. However, we agree with the Associate Editor that regrouping all the data and information on a single web page is relevant. So, we uploaded the source code to the first author
website (\url{http://stevetonneau.fr/files/publications/isrr15/tro_install.html}), and provided all the required information to install the material in it. The user has the choice to download the code either from the github permanent link or directly from the author's website.}

\leavevmode 
\quot {... and to refer
to the underlying method used by the Blender tool.}

\as{Regarding the decimate tool from Blender, we proceeded as follows: in the paper, we mention a SIGGRAPH paper as one option to reduce mesh complexity, and mention that equivalent methods are proposed by many 3D authoring tools, such as Blender which we used. Then, in the static web page, we remind this piece of information and provide a permanent link to the version of Blender we used, along with the documentation that explains how to use the decimate tool.}

\qt{The resulting polytope can contain a very large number of faces.  A last step is thus to simplify it using an incremental decimation method[38].
Variations of this method are commonly implemented in most authoring tools. For our experiments we used the Blender decimate tool. Details of its use can be found in the static webpage associated with this paper \url{http://stevetonneau.fr/files/publications/isrr15/tro_install.html\#decimate}.\\}

@@

\as{Thanks for the comment. Indeed, using the dual formulation is quite common for LP solvers. However, after computing the dual in our case, it turns out that some of the constraints can be solved trivially, which allows to reduce the dimensionality of the problem, resulting in (11). We experimentally verified that this form is faster to solve with respect to the one from equation (10). For this reason, we believe that equation (11) is worth mentioning, and prefer to leave it, since it was given as an option by the reviewer. Finally, we removed the reference to the Slater's conditions since we agree with the reviewer that the procedure is standard.}

\leavevmode \\
\section{Answers to Reviewer 2}
\quot{All the issues I raised in the first review have been addressed in
the revised version. I'm satisfied with most of the answers and the
changes made to the paper. In particular the novelty of the paper with
respect to the conference works is clarified. I appreciate also the new
structure of the manuscript that integrates material previously
relegated to the Appendices into the main body. Now, this work seems to
be self contained enough for a journal. }
\as{Thanks for the comment. We agree with the reviewer that the paper is better this way.}

\leavevmode \\
\quot{
1. I understand the point of view of the authors who focus on
multi-contact planning and would like to refer only briefly to previous
work in motion planning for multi-legged, statically-stable robots. In
principle that is OK, but I suggest to extend the paragraph added at
the end of Section I.A by few sentences that clarify what are the
advantages of your motion planning system with respect to these older
approaches. You are focusing on humanoids, for which these gains are
pretty obvious, but what about multi-legged robots? As an inspiration I
recommend this youtube video showing a simulated hexapod that does
something similar to your HyQ simulated example
\url{https://www.youtube.com/watch?v=Xio1TbjAK80}}

\as{Thanks for the comment. We added the following paragraph in the Section.}
\qt{The following state of the art focuses on contributions proposed for humanoid robots, although our method is also demonstrated on quadruped robots such as HyQ, for which related work also exists [6]. With robots using more than two legs for locomotion, different gait modes can be used to cross cluttered environments, allowing them to remain in a cyclic context. In these works collision avoidance is often treated as a height issue (assuming that all the obstacles are on the ``ground'', and so are the contacts, which prevents from going under a table, or using a wall as contact location for instance). While these approaches are efficient in many cases, we focus on the most generic case, where obstacles are not only on the ground, and cyclic locomotion might not be a solution.}
\leavevmode \\

\quot{
While the answer about the environment model is satisfying, I think
that you should no only refer to the FCL library in a footnote, but to
cite the paper that explains how it works.
}
\as{Thanks for the suggestion. The citation has indeed been included, as reference [32]}
\leavevmode \\

### 2017-12-13 body/intro.tex
r1 - comment 1. remove optimization converges locally
r2 quadruped ? finished ?

\subsection{State of the art}

@@

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorics. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial aspect of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.~\cite{dai2014whole} extended the work of Posa et al.~\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

>>>>

\subsection{State of the art}
The following state of the art focuses on contributions proposed for humanoid robots, although our method is also demonstrated on quadruped robots such as HyQ, for which related work also exists~\cite{BELTER20116918}. With robots using more than two legs for locomotion, different gait modes can be used to cross cluttered environments, allowing them to remain in a cyclic context. In these works collision avoidance is often treated as a height issue (assuming that all the obstacles are on the ``ground'', and so are the contacts, which prevents from going under a table, or using a wall as contact location for instance). While these approaches are efficient in many cases, we focus on the most generic case, where obstacles are not only on the ground, and cyclic locomotion might not be a solution.

@@

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorics. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial aspect of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.~\cite{dai2014whole} extended the work of Posa et al.~\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
